---
title: "EBH_Plotting_Jm3101_v1.0_and_Walnut_2.0_ref"
author: "Houston Saxe"
date: "2/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load neccessary libraries

```{r message=FALSE, , results='hide', include=FALSE}
pacman::p_load(dplyr, tibble, readr, stringr, data.table, ggplot2, sjplot, tidyr)

## Helpful function for quick viewing of wide data
left = function(x, n = 10){head(x[,1:n])}
```

# Load data and annotation
```{r}
dat = fread('DGEresults/CPM_20_filtered.csv') %>% 
  rename(GeneID = V1)

head(dat)

anno = fread('DGEresults/annotation_combined.csv')
```

# How many genes are being expressed from each haplotype?
```{r}
expressed = dat %>% 
  as.data.frame() %>% 
  mutate(GeneID = as.numeric(GeneID)) %>% 
  select(GeneID) %>% 
  left_join(anno) %>%
  distinct(GeneID, Parent_haplotype) %>% 
  count(Parent_haplotype)

p = ggplot(expressed, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  labs(y='Expressed genes')

p

```

# Read in CG results
```{r}
impCG = fread('DGEresults/Limma_results_table_CG.csv')
```

## Total CG DEGs by haplotype
```{r}
ASEsums = impCG %>% distinct(GeneID, Parent_haplotype) %>% count(Parent_haplotype)

p1 = ggplot(ASEsums, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  theme(legend.position = 'none')+
  labs(y='DEGs')+
  ggtitle('Total CG DEGs')

```

# CG positive genes EBH
```{r}
impCG_pos = impCG %>%
  filter(logFC > 0) %>%
  distinct(GeneID, Parent_haplotype) %>% 
  count(Parent_haplotype)

sum(impCG_pos$n)

p2 = ggplot(impCG_pos, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  theme(legend.position = 'none')+
  labs(y='DEGs')+
  ggtitle('CG DEGs (+)')

```

# CG negative genes EBH
```{r}
impCG_neg = impCG %>%
  filter(logFC < 0) %>%
  distinct(GeneID, Parent_haplotype) %>% 
  count(Parent_haplotype)

p3 = ggplot(impCG_neg, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  theme(legend.position = 'none')+
  labs(y='DEGs')+
  ggtitle('CG DEGs (-)')

```
# Put three CG plots together
```{r}
a = ggarrange(p1, p2, p3, nrow = 3)

a
```


# Read in PHY results
```{r}
impPHY = fread('DGEresults/Limma_results_table_PHY.csv')
```

## Total PHY DEGs by haplotype
```{r}
ASEsums = impPHY %>% distinct(GeneID, Parent_haplotype) %>% count(Parent_haplotype)

p1 = ggplot(ASEsums, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  theme(legend.position = 'none')+
  labs(y='DEGs')+
  ggtitle('Total PHY DEGs')

```

# PHY positive genes EBH
```{r}
impPHY_pos = impPHY %>%
  filter(logFC > 0) %>%
  distinct(GeneID, Parent_haplotype) %>% 
  count(Parent_haplotype)

p2 = ggplot(impPHY_pos, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  theme(legend.position = 'none')+
  labs(y='DEGs')+
  ggtitle('PHY DEGs (+)')

```

# PHY negative genes EBH
```{r}
impPHY_neg = impPHY %>%
  filter(logFC < 0) %>%
  distinct(GeneID, Parent_haplotype) %>% 
  count(Parent_haplotype)

p3 = ggplot(impPHY_neg, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  theme(legend.position = 'none')+
  labs(y='DEGs')+
  ggtitle('PHY DEGs (-)')

```

# Put three PHY plots together
```{r}
b = ggarrange(p1, p2, p3, nrow = 3)

b
```

# Read in NEM results
```{r}
impNEM = fread('DGEresults/Limma_results_table_NEM.csv')
```

## Total NEM DEGs by haplotype
```{r}
ASEsums = impNEM %>% distinct(GeneID, Parent_haplotype) %>% count(Parent_haplotype)

p1 = ggplot(ASEsums, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  theme(legend.position = 'none')+
  labs(y='DEGs')+
  ggtitle('Total NEM DEGs')

```

# NEM positive genes EBH
```{r}
impNEM_pos = impNEM %>%
  filter(logFC > 0) %>%
  distinct(GeneID, Parent_haplotype) %>% 
  count(Parent_haplotype)

p2 = ggplot(impNEM_pos, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  theme(legend.position = 'none')+
  labs(y='DEGs')+
  ggtitle('NEM DEGs (+)')

```

# NEM negative genes EBH
```{r}
impNEM_neg = impNEM %>%
  filter(logFC < 0) %>%
  distinct(GeneID, Parent_haplotype) %>% 
  count(Parent_haplotype)

p3 = ggplot(impNEM_neg, aes(Parent_haplotype, n, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  theme(legend.position = 'none')+
  labs(y='DEGs')+
  ggtitle('NEM DEGs (-)')

```

# Put three NEM plots together
```{r}
c = ggarrange(p1, p2, p3, nrow = 3)

c
```
# Put CG, PHY, and NEM figs into one fig
```{r}
f = ggarrange(a, b, c, labels = c('A)', 'B)', 'C)'), nrow = 1)

f

sjPlot::save_plot('DGEresults/CG_PHY_NEM_EBH.png', f, height = 17, width = 20)
```

