---
title: "EBH_Plotting_Jm3101_v1.0_and_Walnut_2.0_ref"
author: "Houston Saxe"
date: "2/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load neccessary libraries
```{r message=FALSE, , results='hide', include=FALSE}
pacman::p_load(dplyr,
               tibble,
               readr,
               stringr,
               data.table,
               ggplot2,
               sjPlot, 
               tidyr,
               ggpubr,
               ggmosaic)

```

# Load data and annotation
```{r}
dat = fread('DGEresults/DGE_CPM_data.csv')

head(dat)

dat$GeneID %>% unique() %>% length()

anno = fread('DGEresults/annotation_combined.csv')
```

# How many genes are being expressed from each haplotype?
```{r}
expressed = dat %>% 
  as.data.frame() %>% 
  mutate(GeneID = as.numeric(GeneID)) %>% 
  select(GeneID) %>% 
  left_join(anno) %>%
  distinct(GeneID, Parent_haplotype) %>% 
  count(Parent_haplotype, name = 'Expressed_gene_count')

p = ggplot(expressed, aes(Parent_haplotype, Expressed_gene_count, fill = Parent_haplotype))+
  geom_col(color = 'black')+
  labs(y='Expressed genes')

p

```

### View table
```{r}
expressed
```

# CG Read in results
```{r}
impCG = fread('DGEresults/Limma_results_table_CG.csv') %>% 
  mutate(Gene_type = ifelse(logFC > 0, 'Susceptibility', 'Resistance'),
         Parent_haplotype = factor(Parent_haplotype, levels = c('J.regia', 'J.microcarpa')))

length(unique(impCG$GeneID))
```

### Total CG DEGs by haplotype
```{r}
ASEsums = impCG %>% 
  distinct(GeneID, 
           Parent_haplotype,
           Gene_type) %>% 
  count(Parent_haplotype,
        Gene_type,
        name = 'DEG_gene_count') %>% 
  pivot_wider(names_from = Parent_haplotype, values_from = DEG_gene_count) %>% 
  column_to_rownames(var = 'Gene_type') %>% 
  select(J.microcarpa, J.regia)

ASEsums

```

### Check to see if sum is correct, should match length(unique(impCG$GeneID))
```{r}
sum(ASEsums)
```

### Fisher's exact test, new structure
```{r}
fisher_res = fisher.test(ASEsums)

fisher_res

```

### New mosaic plot
```{r}
uniq_impCG = impCG %>% 
  distinct(GeneID, 
           Parent_haplotype, 
           Gene_type) %>% 
  mutate(Pathogen = 'A. tumefaciens',
         Stats = paste('Fisher p-value: ',
                     formatC(fisher_res$p.val, format = 'e'),
                     '\n', 'Odds ratio: ',
                     round(fisher_res$estimate, 3), sep = ''))

A = ggplot(uniq_impCG)+
  geom_mosaic(aes(x = product(Parent_haplotype),
                  conds = product(Gene_type), 
                  fill = Parent_haplotype),
              color = 'black')+
  theme_grey(base_size = 14)+
    theme(legend.position = 'none',
          strip.text = element_text(face = 'italic'))+
  facet_wrap(~Pathogen+Stats)

A
```

# PHY Read in results
```{r}
impPHY = fread('DGEresults/Limma_results_table_PHY.csv') %>% 
  mutate(Gene_type = ifelse(logFC > 0, 'Susceptiblity', 'Resistance'),
         Parent_haplotype = factor(Parent_haplotype, levels = c('J.regia', 'J.microcarpa')))

length(unique(impPHY$GeneID))
```

### Total PHY DEGs by haplotype
```{r}
ASEsums = impPHY %>% 
  distinct(GeneID, 
           Parent_haplotype,
           Gene_type) %>% 
  count(Parent_haplotype,
        Gene_type,
        name = 'DEG_gene_count') %>% 
  pivot_wider(names_from = Parent_haplotype, values_from = DEG_gene_count) %>% 
  column_to_rownames(var = 'Gene_type') %>% 
  select(J.microcarpa, J.regia)

```

### Check to see if sum is correct, should match length(unique(impPHY$GeneID))
```{r}
sum(ASEsums)
```

### Fisher's exact test, new structure
```{r}
fisher_res = fisher.test(ASEsums)

fisher_res

```

### New mosaic plot
```{r}
uniq_impPHY = impPHY %>% 
  distinct(GeneID, Parent_haplotype, Gene_type) %>% 
  mutate(Pathogen = 'Phytophthora',
         Stats = paste('Fisher p-value: ',
                     formatC(fisher_res$p.val, format = 'e'),
                     '\n', 'Odds ratio: ',
                     round(fisher_res$estimate, 3), sep = ''))

A = ggplot(uniq_impPHY)+
  geom_mosaic(aes(x = product(Parent_haplotype),
                  conds = product(Gene_type), 
                  fill = Parent_haplotype),
              color = 'black')+
  theme_grey(base_size = 14)+
    theme(legend.position = 'none',
          strip.text = element_text(face = 'italic'))+
  facet_wrap(~Pathogen+Stats)

A
```

# NEM_3Y Read in results
```{r}
impNEM_3Y = fread('DGEresults/Limma_results_table_NEM_3Y.csv') %>% 
  mutate(Gene_type = ifelse(logFC > 0, 'Susceptiblity', 'Resistance'),
         Parent_haplotype = factor(Parent_haplotype, levels = c('J.regia', 'J.microcarpa')))

length(unique(impNEM_3Y$GeneID))
```

### Total NEM_3Y DEGs by haplotype
```{r}
ASEsums = impNEM_3Y %>% 
  distinct(GeneID, 
           Parent_haplotype,
           Gene_type) %>% 
  count(Parent_haplotype,
        Gene_type,
        name = 'DEG_gene_count') %>% 
  pivot_wider(names_from = Parent_haplotype, values_from = DEG_gene_count) %>% 
  column_to_rownames(var = 'Gene_type') %>% 
  select(J.microcarpa, J.regia)

```

### Check to see if sum is correct, should match length(unique(impNEM_3Y$GeneID))
```{r}
sum(ASEsums)
```

### Fisher's exact test, new structure
```{r}
fisher_res = fisher.test(ASEsums)

fisher_res

```

### New mosaic plot
```{r}
uniq_impNEM_3Y = impNEM_3Y %>% 
  distinct(GeneID, Parent_haplotype, Gene_type) %>% 
  mutate(Pathogen = 'P. vulnus',
         Stats = paste('Fisher p-value: ',
                     formatC(fisher_res$p.val, format = 'e'),
                     '\n', 'Odds ratio: ',
                     round(fisher_res$estimate, 3), sep = ''))

A = ggplot(uniq_impNEM_3Y)+
  geom_mosaic(aes(x = product(Parent_haplotype),
                  conds = product(Gene_type), 
                  fill = Parent_haplotype),
              color = 'black')+
  theme_grey(base_size = 14)+
    theme(legend.position = 'none',
          strip.text = element_text(face = 'italic'))+
  facet_wrap(~Pathogen+Stats)

A
```

# Length_3Y Read in results
```{r}
impLength_3Y = fread('DGEresults/Limma_results_table_Length_3Y.csv') %>% 
  mutate(Gene_type = ifelse(logFC > 0, 'Long', 'Short'),
         Parent_haplotype = factor(Parent_haplotype, levels = c('J.regia', 'J.microcarpa')))

length(unique(impLength_3Y$GeneID))
```

### Total Length_3Y DEGs by haplotype
```{r}
ASEsums = impLength_3Y %>% 
  distinct(GeneID, 
           Parent_haplotype,
           Gene_type) %>% 
  count(Parent_haplotype,
        Gene_type,
        name = 'DEG_gene_count') %>% 
  pivot_wider(names_from = Parent_haplotype, values_from = DEG_gene_count) %>% 
  column_to_rownames(var = 'Gene_type') %>% 
  select(J.microcarpa, J.regia)

```

### Check to see if sum is correct, should match length(unique(impLength_3Y$GeneID))
```{r}
sum(ASEsums)
```

### Fisher's exact test, new structure
```{r}
fisher_res = fisher.test(ASEsums)

fisher_res

```

### New mosaic plot
```{r}
uniq_impLength_3Y = impLength_3Y %>% 
  distinct(GeneID, Parent_haplotype, Gene_type) %>% 
  mutate(Pathogen = 'Three-year Tree Length',
         Stats = paste('Fisher p-value: ',
                     formatC(fisher_res$p.val, format = 'e'),
                     '\n', 'Odds ratio: ',
                     round(fisher_res$estimate, 3), sep = ''))

A = ggplot(uniq_impLength_3Y)+
  geom_mosaic(aes(x = product(Parent_haplotype),
                  conds = product(Gene_type), 
                  fill = Parent_haplotype),
              color = 'black')+
  theme_grey(base_size = 14)+
    theme(legend.position = 'none',
          strip.text = element_text(face = 'italic'))+
  facet_wrap(~Pathogen+Stats)

A
```

# Final fig for paper
```{r}
all = list(uniq_impCG,
           uniq_impPHY, 
           uniq_impNEM_3Y,
           uniq_impLength_3Y) %>% 
  bind_rows()

D = ggplot(all)+
  geom_mosaic(aes(x = product(Parent_haplotype),
                  conds = product(Gene_type), 
                  fill = Parent_haplotype),
              color = 'black')+
  theme_grey(base_size = 14)+
    theme(legend.position = 'none',
          strip.text = element_text(face = 'italic'))+
  facet_wrap(~Pathogen+Stats, scales = 'free')

D

sjPlot::save_plot('Final_Figs_and_Draft_Manuscript_SCRI_ROOT/Fig. 3.png', 
                  D,
                  height = 20, 
                  width = 18)
```




