---
title: "Wrangling_SCRI_Root_Phenotype_Data_2"
author: "Houston Saxe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE
)
```

# Load packages
```{r}
pacman::p_load(data.table,
               tidyr,
               dplyr,
               tibble,
               naniar,
               corrplot,
               ggpubr,
               ggplot2,
               ggfortify,
               cowplot,
               OmicsAnalyst)
```

# Read in data
```{r}
LL = openxlsx::read.xlsx("SCRI_Phenotype_Analysis/2023.04.18 SCRI Masterlist with AW updates.xlsx",
                         sheet = 1,
                         startRow = 1)

# LL = openxlsx::read.xlsx("SCRI_Phenotype_Analysis/2023.04.06 SCRI Masterlist.xlsx",
#                          sheet = 1,
#                          startRow = 1)

# LL = openxlsx::read.xlsx("SCRI_Phenotype_Analysis/2023.03.29 SCRI Masterlist DK edited.xlsx",
#                          sheet = 1,
#                          startRow = 1)

# LL = openxlsx::read.xlsx("SCRI_Phenotype_Analysis/2023.02.08 CAL Masterlist lab inventory to CAL fr GTB (1).xlsx",
#                          sheet = 1,
#                          startRow = 2)

# LL = openxlsx::read.xlsx("SCRI_Phenotype_Analysis/2023.03.22 SCRI Masterlist cal_AW.xlsx",
#                          sheet = 1,
#                          startRow = 1)

# LL = read.csv('SCRI_Phenotype_Analysis/2023_masterlist.csv')

# LL = openxlsx::read.xlsx("SCRI_Phenotype_Analysis/WalnutClean_FEB_2023.xlsx",
#                          sheet = 1,
#                          startRow = 1)

# nemTest = LL %>%
#   filter(CAL_Wip_ID %in% c('MS1-36', 'MS1-41', 'MS1-56', 'MS1-122', 'STJM4', '29JM 11', 'JMS 12'))

LL = LL[!duplicated(names(LL))]

## Replacing all non-data containing observations with NAs
LL = replace_with_na_all(LL, ~.x %in% c("", ".")) %>% 
  rename_with(~ gsub('\\.|:|-', '_', .x) %>% 
                gsub('__', '_', .) %>% 
                gsub('>', 'Over_', .) %>% 
                gsub('CAL.*Wip.*ID.*', 'CAL_Wip_ID', .))

head(LL)
```

# Averaging CG scores into one column. This assumes all year scoring is important
```{r}
LL = LL %>%
  mutate(across(-c(matches('CAL..WIP.ID.*'),
                   genotype), ~ gsub('\\(.*?\\)', '', .x))) %>% 
  mutate(across(matches("CG_Phenotype|P_cinn|P_cit|RLN|length"), as.numeric)) %>% 
  mutate(CG_Avg = rowMeans(select(., matches("CG.Phenotype")), na.rm = T) %>% 
         ifelse(is.nan(.), NA, .),
         PHY_Avg = rowMeans(select(., matches('P.cinn|P.cit')), na.rm = T) %>% 
           ifelse(is.nan(.), NA, .),
         Female_Parent = as.numeric(Female_Parent)) %>% 
  relocate(CG_Avg, PHY_Avg, .after = `GTB_Notes_on_value_in_Chiape_Trial,_San_Joaquin_Co`) %>%
  select(!matches('P_cinn|P_cit')) %>% 
  select(!matches("DAK|PJB|SG.|Dave.s|Proposed|X2021|GTB"))

nemTest = LL %>% 
  filter(CAL_Wip_ID %like% 'MS1-36|MS1-41|MS1-56|MS1-122|STJM4|29JM 11|JMS 12')

```

# Where are the NAs coming from?
```{r}
## Sums of NAs for all columns
naSums = LL %>%
  summarise_all(list(~sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "NaSums") %>%
  arrange(NaSums)

## plotting shows where the NAs are coming from
ggplot(naSums, aes(x = NaSums, y = reorder(Variable, NaSums)))+
  geom_bar(stat = "identity")
```

# Picking variables
```{r}
# Data with only 31.01 and 31.09 progeny
LL2 = LL %>%
  select(matches('CAL.WIP.ID'), 
         genotype,
         Female_Parent, 
         CG_Avg, 
         PHY_Avg, 
         matches('length|RLN')) %>%
  filter(Female_Parent %in% c('31.01', '31.09'))

# Data with all female parents
LL_all_parents = LL %>% 
  select(matches('CAL.WIP.ID'), 
         genotype,
         Female_Parent, 
         CG_Avg, 
         PHY_Avg, 
         matches('length|RLN')) 
  
  

## Sums of NAs for all columns
naSums = LL2 %>%
  summarise_all(list(~sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "NaSums") %>%
  arrange(NaSums)

## plotting shows where the NAs are coming from
ggplot(naSums, aes(x = NaSums, y = reorder(Variable, NaSums)))+
  geom_bar(stat = "identity")

length(which(complete.cases(LL2) == TRUE))
```
## NEM data
```{r}
NEM_dat = LL_all_parents %>% 
  select(matches('CAL.WIP.ID'), 
         genotype,
         Female_Parent, 
         matches('length|RLN')) %>% 
  rename_with(~ gsub('2Y_RLN', 'RLN_2Y', .x) %>% 
                gsub('3Y_RLN', 'RLN_3Y', .) %>% 
                gsub('2Y_length', 'length_2Y', .) %>% 
                gsub('3Y_length', 'length_3Y', .))

fwrite(NEM_dat, 'SCRI_Phenotype_Analysis/2_3_Year_RLN_Counts_and_tree_lengths.csv')
```

## CG data
```{r}
CG_dat = LL_all_parents %>% 
  select(matches('CAL.WIP.ID'),
         CG_Avg)

fwrite(CG_dat, 'SCRI_Phenotype_Analysis/CG_Avg_Score.csv')
```

## PHY data
```{r}
PHY_dat = LL_all_parents %>% 
  select(matches('CAL.WIP.ID'),
         PHY_Avg) 

fwrite(PHY_dat, 'SCRI_Phenotype_Analysis/PHY_Avg_Score.csv')
```

# Make and save metadata from RNA-seq experiment
```{r}
metadata = NEM_dat %>% 
  left_join(CG_dat) %>% 
  left_join(PHY_dat)

RNA_seq_dat = metadata %>%
  filter(CAL_Wip_ID %in% c('MS1-36', 'MS1-41', 'MS1-56', 'MS1-122', 'STJM4', '29JM 11', 'JMS 12'))

fwrite(RNA_seq_dat,
       'SCRI_Phenotype_Analysis/2023_metadata.csv')
```

# DF of incomplete cases from LL2
```{r}
LL3 = LL2 %>%
  select(!matches('Over_3Y|2Y')) %>% 
  slice(which(complete.cases(.) == TRUE))


fwrite(LL3, 'SCRI_Phenotype_Analysis/SCRI_Phenotype_Data_Complete_Cases.csv')

LL4 = LL2 %>%
  # remove rows with all NAs except CAL_Wip_ID 
  filter(., !rowSums(is.na(.)) == ncol(.)-1) %>% 
  filter(., !rowSums(is.na(.)) == ncol(.))

fwrite(LL4, 'SCRI_Phenotype_Analysis/SCRI_Phenotype_Data_InComplete_Cases.csv')

```


```{r}
LL3 = fread('SCRI_Phenotype_Analysis/SCRI_Phenotype_Data_Complete_Cases.csv')

Elite_classifying = LL3 %>% 
  mutate(Elite = factor(ifelse(`3Y_RLN` <= 75 &
                                CG_Avg <= 2 &
                                PHY_Avg <= 30, 'Elite',
                                'Not Elite')))
```


# How do phenotypes cluster with eachother?
```{r}
p = plot_pca(LL3 %>%
               select(!c(genotype)) %>% 
               column_to_rownames(var = 'CAL_Wip_ID'),
         center = T,
         scale = T)

p$plot

sjPlot::save_plot('SCRI_Phenotype_Analysis/SCRI_Phenotype_PCA_Complete_Cases_Data.png',
          p$plot,
          dpi = 300,
          width = 15)
```

# How do phenotypes cluster with eachother? Log10 transformation of NEM data
```{r}
RLN_transform = LL3 %>% 
  mutate(across(matches('RLN'), ~ log10(.x+1)))
  

p = plot_pca(RLN_transform %>%
               column_to_rownames(var = 'CAL:CAL_Wip_ID'),
         center = F,
         scale = F)

p$plot

# sjPlot::save_plot('SCRI_Phenotype_Analysis/SCRI_Phenotype_PCA_Complete_Cases_Data_RLN_Transform.png',
#           p$plot,
#           dpi = 300,
#           width = 15)
```

# Correlation of data with complete cases across rows
```{r}
test = as.data.frame(as.table(cor(LL3[,-c(1,2)]))) %>% 
  rename(Cor = Freq)

test2 = as.data.frame(as.table(cor.mtest(LL3[,-c(1,2)])$p)) %>% 
  rename(P.val = Freq)

cor_Complete_cases = test %>% 
  left_join(test2) %>% 
  distinct(Cor, P.val,
           .keep_all = T) %>% 
  mutate(FDR = p.adjust(P.val, method = 'BH'))

```

# Correlation of data with incomplete cases across rows (NAs not removes)
```{r}
LT = cor(LL4[,-c(1,2)],
         use = "pairwise.complete.obs",
         method = 'spearman')

LT[lower.tri(LT)] = NA

test = LT %>% 
  as.table() %>% 
  as.data.frame() %>% 
  drop_na() %>% 
  rename(Cor = Freq) %>% 
  filter(Cor != 1)
  

test2 = as.data.frame(as.table(cor.mtest(LL4[,-c(1,2)])$p)) %>% 
  rename(P.val = Freq)

LT = cor.mtest(LL4[,-c(1,2)],
         use = "pairwise.complete.obs",
         method = 'spearman')$p

LT[lower.tri(LT)] = NA

test2 = LT %>% 
  as.table() %>% 
  as.data.frame() %>% 
  drop_na() %>% 
  rename(P.val = Freq) 

Cor_all_cases = test %>% 
  left_join(test2) %>%
  distinct(Cor, P.val,
           .keep_all = T) %>% 
  mutate(FDR = p.adjust(P.val, method = 'BH'))


fwrite(Cor_all_cases,
       'SCRI_Phenotype_Analysis/SCRI_Phenotype_Data_Spearman_Correlation_InComplete_Cases_Table.csv')
```


```{r}
LT = cor(LL4[,-c(1,2)],
         use = "pairwise.complete.obs",
         method = 'spearman') %>% 
  data.frame() 

LT_Noah = LT %>% 
  select(PHY_Avg)

LT_p = cor.mtest(LL4[,-c(1,2)],
         use = "pairwise.complete.obs",
         method = 'spearman')$p

LT_P_Noah = LT_p %>% 
  data.frame() %>% 
  select(PHY_Avg) %>%
  rename(P.val = PHY_Avg) %>% 
  mutate(FDR = p.adjust(P.val)) 

final = LT_P_Noah %>% 
  rownames_to_column(var = 'Vars') %>% 
  left_join(LT_Noah %>% 
              rownames_to_column(var = 'Vars'))
```

